---
title: "Mice project report"
author: "Pavel Vychyk"
date: "10/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Check for required packages
```{r perform packages installation, message=FALSE, results='hide', warn.conflicts=F}
pckages <- c('xlsx', 'ggplot2', 'ggpubr', 'car', 'FSA', 'ggfortify', 'knitr')
for(package in pckages) {
  if(!require(package, character.only = T)) {
    install.packages(package, dependencies = T)
    library(package)
  }
}
```
#### Load required libraries 
```{r message=FALSE}
library(xlsx)
library(ggplot2)
library(ggpubr)
library(car)
library(FSA)
library(ggfortify)
library(knitr)
```

#### provide the path to the data file and read
```{r}
data_mice <- read.xlsx("../data_folder/Data_Cortex_Nuclear.xls", "Hoja1")
```

#### 1. EDA
Lets look on the data. We could see, that column MouseID has repeats for the
same object
```{r}
head(data_mice)
```
Get the number of mice in the data set
```{r}
length(unique(sapply(strsplit(data_mice$MouseID, "_"), "[", 1)))
```
Look on the presented groups
```{r} 
unique(data_mice$class)
unique(data_mice$Treatment)
unique(data_mice$Genotype)
unique(data_mice$Behavior)
```
Check the groups balance
``` {r}
table(data_mice$class)
table(data_mice$Treatment)
table(data_mice$Genotype)
table(data_mice$Behavior)
```
Count NA in samples
```{r}
sapply(data_mice, function(y) sum(is.na(data_mice)))
nrow(data_mice[rowSums(is.na(data_mice))>0, ])
```

#### 2. Is there any difference in BDNF_N expression levels regarding to the class
Let's drop incomplete cases
```{r} 
complete_data <- data_mice[complete.cases(data_mice),]
```
And look on the assumptions for comparison tests,
firstly check normality distribution using graphical analysis
and Shapiro–Wilk test
QQ-plot:
```{r}
ggqqplot(complete_data$BDNF_N)
```
Shapiro-Wilk test result:
```{r}
shtest <- shapiro.test(complete_data$BDNF_N)
shapiro.test(complete_data$BDNF_N)
```
Check homogeneity of variances assumption within classes
```{r}
complete_data$class <- as.factor(complete_data$class)
lvtest <- leveneTest(BDNF_N ~ class, data=complete_data)
leveneTest(BDNF_N ~ class, data=complete_data)
``` 
Despite the QQ-plot shows near normal distribution,
Shapiro-Wilk test <b>rejects data normality</b> hypothesis <b>(p-value = `r shtest$p.value`)</b>
Levene's test <b>rejects variance homogeneity</b> <b>(p-value = `r lvtest$p.value`)</b>.
Due to lack of default assumptions in our data, start with non-parametric Kruskal-Wallis test for difference
between classes.
```{r}
kwtest <- kruskal.test(BDNF_N ~ class, data=complete_data)
kruskal.test(BDNF_N ~ class, data=complete_data)
``` 
Kruskal-Wallis test <b>rejects the null hypothesis</b> for equality of medians within groups 
<b>(p-value = `r kwtest$p.value`)</b>. We should perform post-hoc analysis to
find which groups have significant difference in medians. 
For post-hoc we use Dunn's test:
```{r}
dTest <- dunnTest(BDNF_N ~ class, data=complete_data, method = "bh")
dResult <- dTest$res
dResult[dResult$P.adj < 0.05, ]
```
Dunn test shows statistical significant difference in medians for classes at p-value level < 0.05,
most significant difference was for classes below:
c-SC-m: control mice, not stimulated to learn, injected with memantine
t-SC-s: trisomy mice, not stimulated to learn, injected with saline
Let's visualize BDNF_N expressions difference for classes with boxplot graphs:
```{r}
ggplot(complete_data, aes(x = class, y = BDNF_N, col = class)) + geom_boxplot()
```
#### 3. Try to predict BDNF_N expression based on other proteins expression levels with linear model
Build linear model using as predictor variables all columns, except categorical data
```{r}
f_model <- lm(BDNF_N ~ . - MouseID - Genotype - Treatment - Behavior - class, data = complete_data)
summary(f_model)
```
R-square seems significant, but is our model really good?
Residuals vs Fitted shows good support for the linear assumptions
```{r}
plot(f_model, 1)
```
Normal QQ shows the presence of several outliers, but in general model's residual are normally distributed
```{r}
plot(f_model, 2)
```
Spread location graph shows indication of heteroscedasticity, due to unequal 
```{r}
plot(f_model, 3)
```
Cook's distance also plots some outliers
```{r}
plot(f_model, 4)
```
Multicollinearity within the data could
be checked with correlation matrix, we could
see that many pairs demonstrate high level of correlation (> 0.6) 
```{r}
cor(complete_data[,2:78])
```
We could conclude, that build linear model lacks
basic support, and could not be used for prediction.